Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-15 writing/embedded_meta_learning.tex
Style tweaks + more length

\begin{figure}
\centering
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/meta_results.png}
\caption{From examples of mapping.}
\label{poly_meta_map_results_examples}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\end{subfigure}
\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The solid line indicates optimal performance; the dashed line indicates chance (untrained). Error-bars are bootstrap 95\%-CIs.} 
\label{poly_meta_map_results}
\end{figure}
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results}. \par

\textbf{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of their application or receiving a natural-language instruction. \par

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/continual/continual_learning.png}
\caption{Once the meta-learning system has been trained on a distribution of prior tasks, its performance on new tasks can be tuned by caching its guessed embeddings for the tasks and then optimizing those, thus avoiding any possibility of interfering with performance on prior tasks. Starting from random embeddings in the trained model results in slower convergence, while in an untrained model the embeddings cannot be optimized well. Error-bars are bootstrap 95\%-CIs.}
\label{poly_continual_results}
\end{figure}
\textbf{Continual learning:} Although the meta-learning approach is effective for rapidly adapting to a new task, it is unreasonable to think that our system must consider every example it has seen at each inference step. We would like to be able to store our knowledge more efficiently, and allow for further refinement. Furthermore, we would like the system to be able to adapt to new tasks (for which its guessed solution isn't perfect) without catastrophically interfering with prior tasks \citep{McCloskey1989}. \par
A very simple solution to these problems is naturally suggested by our architecture. Specifically, task embeddings can be cached so that they don't have to be regenerated at each inference step. This also allows optimization of these embeddings without altering the other parameters in the architecture, thus allowing fine-tuning on a task without seeing more examples, and without interfering with performance on any other task. That is, we can see the meta-learning step as a ``warm start'' for an optimization procedure over embeddings that are cached in memory \citep[c.f.][]{Kumaran2016}.\footnote{Note that the weights generated by the hyper network could be cached instead of the embeddings, which would be much more memory intensive but would allow even better tuning and would not require passing through the hyper network at inference time.} \par 
To test this idea, we pre-trained the system on 100 polynomial tasks, and then introduced 100 new tasks. We trained on these new tasks by starting from the meta-network's ``guess'' at the correct task embedding, and then optimizing this embedding without altering the other parameters. The results are shown in fig. \ref{poly_continual_results}. The meta-network embeddings offers good immediate performance, and substantially accelerates the optimization the process, compared to a randomly-initialized embedding (see supp. fig. \ref{supp_poly_continual_results} for a more direct comparison). Furthermore, this ability to learn is due to training, not simply the expressiveness of the architecture, as is shown by attempting the same with an untrained network. \par

@@

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of playing a held-out game from examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. Note that the sample complexity in terms of number of trained tasks is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. Finally, the system can integrate new tasks rapidly (fig. \ref{cards_basic_integration_results}).\par
\textbf{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior. For example, if the system is trained to map games to their losers variations, it can generalize this mapping to a game it has not been trained to map, even if the source or target of that mapping is held out from training. In Fig. \ref{cards_meta_map_results}) we demonstrate this by taking the mapped embedding and evaluating the reward received by playing the targeted game with it. This task is more difficult than simply learning to play a held out game from examples, because the system will actually receive no examples of the target game (when it is held out). Furthermore, in the case of the losers mapping, leaving the strategy unchanged would produce a large negative reward, and chance performance would produce 0 reward, so the results are quite good. \par

@@

\textbf{Related work:} Our work is an extrapolation from the rapidly-growing literature on meta-learning \citep[e.g.][]{Vinyals2016, Santoro2016, Finn2017a, Finn2018, Stadie2018}. It is also related to the literature on continual learning, or more generally tools for avoiding catastrophic interference based on changes to the architecture \citep[e.g.][]{Fernando2017, Rusu2016}, loss \citep[e.g.][]{Kirkpatrick2016, Zenke2017, Aljundi2019}, or external memory \citep[e.g.][]{Sprechmann2018}. Recent work has also begun to blur the separation between these approaches, for example by meta-learning in an online setting \citep{Finn2019}. Our work is specifically inspired by the algorithms that attempt to have the system learn to adapt to a new task via activations rather than weight updates, either from examples \citep[e.g.][]{Wang2016a, Duan2016}, or a task input \citep[e.g.][]{Borsa2019}. \par

@@

\textbf{Future Directions:} We think that the general perspective of embedding tasks and considering meta-mappings will yield many fruitful future directions. We hope that our work will inspire more exploration of behavioral adaptation in areas beyond the simple domains we considered here. To this end, we suggest the creation of meta-learning datasets which include information not only about tasks, but about the relationships between them. For example, as we alluded to above, tasks such as visual question-answering \citep[e.g.][]{Antol2015} can be usefully interpreted from this perspective (a question essentially yields a function from images to answers, and there are many systematic relationships between these functions), as can reinforcement learning tasks which involve executing instructions \citep[e.g.][]{Hermann2017, Co-Reyes2019}. Furthermore, we think our work provides a novel perspective on the types of flexibility that human intelligence exhibits, and thus hope that it may have implications for cognitive science. \par 

@@

\textbf{Conclusions:} In this paper, we've highlighted a new type of flexibility in the form of \emph{meta-mappings}, that is, mappings between tasks. We've suggested that achieving this type of flexibility is difficult in most deep learning architectures, because they separate computation from the knowledge of what is being computed. We've proposed that representing the task explicitly as the computation provides a solution to this problem. To this end, we've proposed a meta-learning system which embeds tasks and data into a shared latent space, and then meta-learns to make mappings over this space. The system can do this with quite small meta-task sample complexity, since it can generalize to unseen meta-mappings after training on only 20 examples.\par
We see our proposal as a logical progression from the fundamental idea of meta-learning -- that there is a continuum between data and tasks. This naturally suggests the idea of manipulating task representations just like we manipulate data. We've shown that this approach yields considerable flexibility, most importantly the meta-mapping ability to adapt zero-shot to a new task, from examples of task transformations. We've also shown that our approach provides a useful perspective on continual learning. We hope that these results will lead to the development of more powerful and flexible deep-learning models. \par

>>>>

\includegraphics[width=\textwidth]{figures/poly/meta_results.png}
\label{poly_meta_map_results_examples}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\label{poly_meta_map_results_language}
\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The solid line indicates optimal performance; the dashed line indicates chance (untrained). Error-bars are bootstrap 95\%-CIs.} 
\label{poly_meta_map_results}
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par
\vspace{-0.7em}
\paragraph{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results}. \par 
\vspace{-0.7em}
\paragraph{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of their application or receiving a natural-language instruction. \par 
\includegraphics[width=\textwidth]{figures/continual/continual_learning.png}
\caption{Once the meta-learning system has been trained on a distribution of prior tasks, its performance on new tasks can be tuned by caching its guessed embeddings for the tasks and then optimizing those, thus avoiding any possibility of interfering with performance on prior tasks. Starting from random embeddings in the trained model results in slower convergence, while in an untrained model the embeddings cannot be optimized well. Error-bars are bootstrap 95\%-CIs.}
\label{poly_continual_results}
\vspace{-0.7em}
\paragraph{Continual learning:} Although the meta-learning approach is effective for rapidly adapting to a new task, it is unreasonable to think that our system must consider every example it has seen at each inference step. We would like to be able to store our knowledge more efficiently, and allow for further refinement. Furthermore, we would like the system to be able to adapt to new tasks (for which its guessed solution isn't perfect) without catastrophically interfering with prior tasks \citep{McCloskey1989}. \par
A very simple solution to these problems is naturally suggested by our architecture. Specifically, task embeddings can be cached so that they don't have to be regenerated at each inference step. This also allows optimization of these embeddings without altering the other parameters in the architecture, thus allowing fine-tuning on a task without seeing more examples, and without interfering with performance on any other task. That is, we can see the meta-learning step as a ``warm start'' for an optimization procedure over embeddings that are cached in memory \citep[c.f.][]{Kumaran2016}.\footnote{Note that the weights generated by the hyper network could be cached instead of the embeddings, which would be much more memory intensive but would allow even better tuning and would not require passing through the hyper network at inference time.} \par 
To test this idea, we pre-trained the system on 100 polynomial tasks, and then introduced 100 new tasks. We trained on these new tasks by starting from the meta-network's ``guess'' at the correct task embedding, and then optimizing this embedding without altering the other parameters. The results are shown in fig. \ref{poly_continual_results}. The meta-network embeddings offer good immediate performance, and substantially accelerate the optimization process, compared to a randomly-initialized embedding (see supp. fig. \ref{supp_poly_continual_results} for a more direct comparison). Furthermore, this ability to learn is due to training, not simply the expressiveness of the architecture, as is shown by attempting the same with an untrained network. \par

@@

\vspace{-0.7em}
\paragraph{Basic meta learning:} First, we show that the system is able to achieve the basic goal of playing a held-out game from examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. Note that the sample complexity in terms of number of trained tasks is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. Finally, the system can integrate new tasks rapidly (fig. \ref{cards_basic_integration_results}).\par
\vspace{-0.7em}
\paragraph{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior. For example, if the system is trained to map games to their losers variations, it can generalize this mapping to a game it has not been trained to map, even if the source or target of that mapping is held out from training. In Fig. \ref{cards_meta_map_results}) we demonstrate this by taking the mapped embedding and evaluating the reward received by playing the targeted game with it. This task is more difficult than simply learning to play a held out game from examples, because the system will actually receive no examples of the target game (when it is held out). Furthermore, in the case of the losers mapping, leaving the strategy unchanged would produce a large negative reward, and chance performance would produce 0 reward, so the results are quite good. \par

@@

\paragraph{Related work:} Our work is an extrapolation from the rapidly-growing literature on meta-learning \citep[e.g.][]{Vinyals2016, Santoro2016, Finn2017a, Finn2018, Stadie2018}. It is also related to the literature on continual learning, or more generally tools for avoiding catastrophic interference based on changes to the architecture \citep[e.g.][]{Fernando2017, Rusu2016}, loss \citep[e.g.][]{Kirkpatrick2016, Zenke2017, Aljundi2019}, or external memory \citep[e.g.][]{Sprechmann2018}. Recent work has also begun to blur the separation between these approaches, for example by meta-learning in an online setting \citep{Finn2019}. Our work is specifically inspired by the algorithms that attempt to have the system learn to adapt to a new task via activations rather than weight updates, either from examples \citep[e.g.][]{Wang2016a, Duan2016}, or a task input \citep[e.g.][]{Borsa2019}. \par

@@

\vspace{-1em}
\paragraph{Future Directions:} We think that the general perspective of embedding tasks and considering meta-mappings will yield many fruitful future directions. We hope that our work will inspire more exploration of behavioral adaptation in areas beyond the simple domains we considered here. To this end, we suggest the creation of meta-learning datasets which include information not only about tasks, but about the relationships between them. For example, as we alluded to above, tasks such as visual question-answering \citep[e.g.][]{Antol2015} can be usefully interpreted from this perspective (a question essentially yields a function from images to answers, and there are many systematic relationships between these functions), as can reinforcement learning tasks which involve executing instructions \citep[e.g.][]{Hermann2017, Co-Reyes2019}. Furthermore, we think our work provides a novel perspective on the types of flexibility that human intelligence exhibits, and thus hope that it may have implications for cognitive science. \par 

@@

\vspace{-1em}
\paragraph{Conclusions:} In this paper, we've highlighted a new type of flexibility in the form of \emph{meta-mappings}, that is, mappings between tasks. We've suggested that achieving this type of flexibility is difficult in most deep learning architectures, because they separate computation from the knowledge of what is being computed. We've proposed that explicitly representing the tasks by embeddings, and deriving the computation from these embeddings, provides a solution to this problem. To this end, we've proposed a meta-learning system which embeds tasks and data into a shared latent space, and then meta-learns to make mappings over this space. The requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings.\par
We see our proposal as a logical progression from the fundamental idea of meta-learning -- that there is a continuum between data and tasks. This naturally leads to the idea of manipulating task representations just like we manipulate data. We've shown that this approach yields considerable flexibility, most importantly the meta-mapping ability to adapt zero-shot to a new task. We've also shown that our approach provides a useful perspective on continual learning. We hope that these results will lead to the development of more powerful and flexible deep-learning models. \par
